---
title: "Coursera Statistical Inference Part 1"
author: "Stella"
date: '2022-12-09'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project Summary

This is part one to the two part projects (1. A simulation excercise, 2. Basic inferential data analysis). Here, I will investigate the exponential distribution in R and compare it to the Central Limit Theorem. The mean and standard deviation of the exponential distribution is 1/lambda. Lambda will be set to 0.2 for all simulations, and there will be 1000 simulations. I will also determine the distribution of averages for 40 exponentials. 


## Simulations
Set the seed, the rate parameter, the number of exponentials, and the number of simulations based on the rubric:
```{r simulation, echo=FALSE}
set.seed(1234)

lambda <- 0.2
n <- 40
sim <- 1000
```
simulate the 40 exponentials, then calculate the mean.
```{r simulationforty, echo=FALSE}
simExp <- replicate(sim, rexp(n, lambda))
meanExp <- apply(simExp, 2, mean)
```
## Answering Question 1
Calculate and compare the sample mean to the theoretical mean.
```{r questionOne, echo=FALSE}
sampMean <- mean(meanExp)
sampMean
theoMean <- 1/lambda
theoMean
```
Develop a histogram of the simulated exponential sample means. The red indicates the sample mean while the blue indicates the theoretical mean.
```{r questionOneGraph, echo=FALSE}
hist(meanExp, main = "Simulated Exponential Sample Means", col = "aliceblue", breaks = 100)
abline(v = sampMean, col = "red")
abline(v = theoMean, col = "blue")
```
Based on the graph, the sample mean is very close to the theoretical mean. 

## Answering Question 2
Calculate and compare the sample variance to the theoretical variace. 
```{r questionTwo, echo=FALSE}
sampSd <- sd(meanExp)
sampVar <- sampSd^2
sampVar

theoSd <- (1/lambda)/sqrt(n)
theoVar <- theoSd^2
theoVar
```
The sample variance is equal to 0.571 which is very close to the theoretical variance of 0.625.

## Answering Question 3
Investigate whether the exponential distribution is approximately normal. Based on CLT, the mean should be relatively normal.
```{r questionThree, echo=FALSE}
hist(meanExp, main = "Normal Distribution", col = "aliceblue", breaks = 100)
xfit <- seq(min(meanExp), max(meanExp), length = 100)
yfit <- dnorm(xfit, mean = 1/lambda, sd = (1/lambda)/sqrt(n))
lines(xfit, yfit*60, lty = 5)
```
Based on the figure, the distribution of means follows a normal curve relatively closely. If there were to be more trials (>1000) the curve would likely be closer to a normal distribution. 